---
title: "Homework 3"
author: "Suning Zhao"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document

---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


### Problem 1

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

### Problem 2

#### Read in the data, tidy and wrangle

* The process begins with data import and updating variable names, using `read_csv` and `clean_names()`.
* Then, I create a new variable called `weekdays` using `case_when()` to classify whether the days are in weekend or weekdays.
* In the next step, I use `pivot_longer` to make it a longer table for accelerometer in different minutes, and then make `minutes` as numeric
* Finally, I change the order of columns using `fct_level`,`select` and arrange the dataset using `arrange`.

```{r}
Accel_df =
  read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    weekdays = case_when(
      day %in% c("Saturday","Sunday") ~ "Weekend",
      day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "Weekday",
      TRUE ~ "")
      )%>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minutes",
    names_prefix = "activity_",
    values_to = "accelerometer"
  ) %>%
  mutate(
    day = forcats::fct_relevel(day, "Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday"),
    minutes = as.numeric(minutes)
  ) %>% 
  select(week, day_id, weekdays, day, everything()) %>% 
  arrange(week, day)
```

* In the dataset `Accel_df`, there are `r nrow(Accel_df)` observations and `r ncol(Accel_df)` variables. 
* The key variables' names are `r names(Accel_df)`.

#### Traditional analyses of accelerometer data

* To analyze the accelerometer data focusing on the total activity over the day, I used `group_by()` and summarize()` to generate a total activity for each day.
* Then, I use `pivot_wider` to generate a wider table to make it clearer, and use `kable` in `knitr` to create the visualized table

```{r}
Accel_df %>% 
  group_by(week,day) %>% 
  summarize(
    sum_accelerometer = sum(accelerometer, na.rm = TRUE)
  ) %>% 
  pivot_wider(
    names_from = "day",
    values_from = "sum_accelerometer"
  ) %>% 
  knitr::kable(digits = 3)
```

* The result from the table seems that there is no apparent trends for total activity in each day, while I notice that it is weird to find that on Saturday in week 4 and week 5, there is only 1440 activity counts there.

#### Make a single-panel plot

* For this question, I use `ggplot` to make x-axis as `minutes` and y-axis as `accelerometer` and use color to indicate the day of the week.

```{r}
Accel_plot = 
  Accel_df %>% 
  group_by(day_id) %>% 
  ggplot(aes(x = minutes, y = accelerometer, color = day)) +
  geom_line(alpha = .5)+
  labs(
    title = "Inspection activity over the course of the day",
    x = "Time of the day",
    y = "Activity counts" ) +
  scale_x_continuous(
    breaks = c(0, 180, 360, 540, 720, 900, 1080, 1260, 1440),
    labels = c("0:00", "3:00", "6:00", "9:00", "12:00", "15:00", "18:00", "21:00", "24:00")
  )

ggsave("./Results/Accel_plot.pdf", Accel_plot, width = 8, height = 5)

Accel_plot
```

* From the graph we can know that:
  1. Within one day, the patient is with low activity level in the first 6hr, and with a relatively higher activity from 6:00 to 12:00 and from 18:00 to 22:00.
  2. According to the paragraph, for one patient, the activity level is relatively higher on Monday, Tuesday and Wednesday. 

### Problem 3

#### Read in the data and do some exploration

```{r}
library(p8105.datasets)
data("ny_noaa")
summary(ny_noaa)
```

* In the original dataset `ny_noaa`, there are `r nrow(ny_noaa)` observations and `r ncol(ny_noaa)` variables. 
* The key variables' names are `r names(ny_noaa)`.
* There are lots of missing data in the dataset. There are `r sum(is.na(ny_noaa$prcp))` missing value in `prcp`, `r sum(is.na(ny_noaa$snow))` missing value in `snow`, `r sum(is.na(ny_noaa$snwd))` missing value in `snwd`. 
* Besides, due to the missing value in tmax and tmin, those two variables are character variables and might need some change.

#### Data cleaning

* I create a new dataframe called `ny_noaa_tidy` to clean the data.
* First, I use `clean_names()` to make sure every variables are with same format. 
* Then, I use `separate()` to separate `date` into `year`, `month`, and `day`.
* Next, I convert `tmin` and `tmax` into numeric variables, and translate `tmin`, `tmax` into "degrees C"; `prcp` into "mm".
* finally, I count the `snow` variable to get the most commonly observed values.
```{r}
ny_noaa_tidy =
  ny_noaa %>% 
  janitor::clean_names() %>% 
  separate(col = date, into = c('year','month','day'), sep = "-" , convert = TRUE) %>% 
  mutate(
     month = month.name[month],
     prcp = prcp / 10,
     tmax = as.numeric(tmax),
     tmin = as.numeric(tmin),
     tmax = tmax / 10,
     tmin = tmin / 10
  ) 
  

ny_noaa_tidy %>%
  count(snow, name = "n_obs") %>% 
  arrange(desc(n_obs))
```

* The result shows that "0" is the most commonly observed values, because in most of months and days, there is no snow in New York state. This aligns with our common sense. 

#### Two-panel plot

```{r}
ny_noaa_max_temp_plot = 
  ny_noaa_tidy %>% 
  filter(month == c("January","July")) %>% 
  group_by(id, year, month) %>% 
  summarize(
    tmax_average = mean(tmax, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = year, y = tmax_average, color = id)) +
  geom_point(alpha = .5) +
  theme(legend.position = "none") +
  facet_grid(. ~ month) +
  labs(
    title = "Average max temperature in January and in July for each station across years",
    y = "Average max temperature (degree C)" ) 

ggsave("./Results/ny_noaa_max_temp_plot.pdf", ny_noaa_max_temp_plot, width = 8, height = 5)

ny_noaa_max_temp_plot
```



